---
title: "Mini Data Analysis Milestone 2"
output: github_document
---

*To complete this milestone, you can either edit [this `.rmd` file](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/mini-project/mini-project-2.Rmd) directly. Fill in the sections that are commented out with `<!--- start your work here--->`. When you are done, make sure to knit to an `.md` file by changing the output in the YAML header to `github_document`, before submitting a tagged release on canvas.*

# Welcome to the rest of your mini data analysis project!

In Milestone 1, you explored your data. and came up with research questions. This time, we will finish up our mini data analysis and obtain results for your data by: 

- Making summary tables and graphs 
- Manipulating special data types in R: factors and/or dates and times.
-   Fitting a model object to your data, and extract a result.
-   Reading and writing data as separate files.

We will also explore more in depth the concept of *tidy data.* 

**NOTE**: The main purpose of the mini data analysis is to integrate what you learn in class in an analysis. Although each milestone provides a framework for you to conduct your analysis, it's possible that you might find the instructions too rigid for your data set. If this is the case, you may deviate from the instructions -- just make sure you're demonstrating a wide range of tools and techniques taught in this class.

# Instructions

**To complete this milestone**, edit [this very `.Rmd` file](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/mini-project/mini-project-2.Rmd) directly. Fill in the sections that are tagged with `<!--- start your work here--->`.

**To submit this milestone**, make sure to knit this `.Rmd` file to an `.md` file by changing the YAML output settings from `output: html_document` to `output: github_document`. Commit and push all of your work to your mini-analysis GitHub repository, and tag a release on GitHub. Then, submit a link to your tagged release on canvas.

**Points**: This milestone is worth 50 points: 45 for your analysis, and 5 for overall reproducibility, cleanliness, and coherence of the Github submission. 

**Research Questions**: In Milestone 1, you chose two research questions to focus on. Wherever realistic, your work in this milestone should relate to these research questions whenever we ask for justification behind your work. In the case that some tasks in this milestone don't align well with one of your research questions, feel free to discuss your results in the context of a different research question.

# Learning Objectives

By the end of this milestone, you should:

-   Understand what *tidy* data is, and how to create it using `tidyr`.
-   Generate a reproducible and clear report using R Markdown.
-   Manipulating special data types in R: factors and/or dates and times.
-   Fitting a model object to your data, and extract a result.
-   Reading and writing data as separate files.

# Setup

Begin by loading your data and the tidyverse package below:

```{r, message = FALSE}
library(datateachr) # <- might contain the data you picked!
library(tidyverse)
```

# Task 1: Process and summarize your data 

From milestone 1, you should have an idea of the basic structure of your dataset (e.g. number of rows and columns, class types, etc.). Here, we will start investigating your data more in-depth using various data manipulation functions. 

### 1.1 (1 point) 

First, write out the 4 research questions you defined in milestone 1 were. This will guide your work through milestone 2:

<!-------------------------- Start your work below ---------------------------->
1. How does the diagnosis of breast cancer vary by the size and shape of the tumor cells?

2. What are the most important predictors of breast cancer diagnosis among the variables in the dataset?

3. How does the diagnosis of breast cancer vary by the age and stage of the patients?

4. How does the diagnosis of breast cancer vary by the smoothness and concavity of the tumor cells?
<!----------------------------------------------------------------------------->

Here, we will investigate your data using various data manipulation and graphing functions.

### 1.2 (8 points)

Now, for each of your four research questions, choose one task from options 1-4 (summarizing), and one other task from 4-8 (graphing). You should have 2 tasks done for each research question (8 total). Make sure it makes sense to do them! (e.g. don't use a numerical variables for a task that needs a categorical variable.). Comment on why each task helps (or doesn't!) answer the corresponding research question.

Ensure that the output of each operation is printed!

Also make sure that you're using dplyr and ggplot2 rather than base R. Outside of this project, you may find that you prefer using base R functions for certain tasks, and that's just fine! But part of this project is for you to practice the tools we learned in class, which is dplyr and ggplot2.

**Summarizing:**

1.  Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data.
2.  Compute the number of observations for at least one of your categorical variables. Do not use the function `table()`!
3.  Create a categorical variable with 3 or more groups from an existing numerical variable. You can use this new variable in the other tasks! *An example: age in years into "child, teen, adult, senior".*
4. Compute the proportion and counts in each category of one categorical variable across the groups of another categorical variable from your data. Do not use the function `table()`!

**Graphing:**

6. Create a graph of your choosing, make one of the axes logarithmic, and format the axes labels so that they are "pretty" or easier to read.
7. Make a graph where it makes sense to customize the alpha transparency.

Using variables and/or tables you made in one of the "Summarizing" tasks: 

8. Create a graph that has at least two geom layers. 
9. Create 3 histograms, with each histogram having different sized bins. Pick the "best" one and explain why it is the best.

Make sure it's clear what research question you are doing each operation for!

<!------------------------- Start your work below ----------------------------->

# 1. How does the diagnosis of breast cancer vary by the size and shape of the tumor cells?

In order to determine how diagnosis varies by the size and shape of the tumor cells, we use the `cancer_sample` dataset in R, which contains information about 569 patients with breast cancer. We compute some statistical parameters such as range, mean, median and standard deviation of two numerical variables: `area_se`, which is the standard error of the area of the tumor cells in square millimeters, and `concavity_worst`, which is the worst or largest concavity of the tumor cells (dimensionless). We compare these parameters across the `diagnosis` variable, which is a categorical variable that indicates whether the tumor is benign or malignant. The output will help us to understand how these variables influence the diagnosis.

```{r}
cancer_sample %>% 
  group_by(diagnosis) %>% 
  summarize(mean_area_se = mean(area_se),
            median_area_se = median(area_se),
            range_area_se = paste(min(area_se), max(area_se), sep = " to "),
            sd_area_se = sd(area_se),
            mean_concavity_worst = mean(concavity_worst),
            median_concavity_worst = median(concavity_worst),
            range_concavity_worst = paste(min(concavity_worst), max(concavity_worst), sep = " to "),
            sd_concavity_worst = sd(concavity_worst)
            )

```
The results show that malignant tumors have higher mean, median, range, and standard deviation of both area_se and concavity_worst than benign tumors. The mean values of `area_se` and `concavity_worst` are 72.67 and 0.45 for malignant tumors, while they are 21.14 and 0.17 for benign tumors. The range of `area_se` for malignant tumors is from 13.99 to 542.2, while for benign tumors it is from 6.802 to 77.11. The range of `concavity_worst` for malignant tumors is from 0 to 1.17, while for benign tumors it is from 0 to 1.252. 
This means that malignant tumors have larger and more variable area and concavity than benign tumors. These variables can help us distinguish between benign and malignant tumors, as larger and more concave cells are more likely to be cancerous than smaller and less concave cells.

Next, we compute the observations in the categorical variable `diagnosis`. 

```{r}
cancer_sample %>% 
  count(diagnosis) %>% 
  mutate(prop = n/sum(n))
```
The output shows that there are 357 benign tumors while there are 212 malignant tumors. The proportion of benign tumors in the data is 0.6274, while the proportion of malignant tumors is 0.3726. This indicates that there are more benign tumors than malignant tumors in the data.

**Visualizing**

In order to understand our data and answer our research question, we will visualize the data using a histogram. A histogram will help us understand the distribution of the numerical variable `area_se` that measures the standard error of the area of tumor cells in square millimeters by the categorical variable `diagnosis`. This will help us compare the size and shape of the tumor cells between benign and malignant types.

```{r}
cancer_sample %>% 
  group_by(diagnosis) %>% 
  ggplot(aes(log(area_se), fill = diagnosis)) +
  geom_histogram(bins = 30, alpha = 0.5, position = "dodge") +
  labs(x = "Log of standard error of area of nuclei \n present in sample image",
       y = "Count",
       title = "Density distribution of standard error of area of nuclei of cancer patients by diagnosis") +
  theme_minimal() +
  scale_fill_manual(values = c("B" = "maroon", "M" = "skyblue"),
                    labels = c("B" = "Benign (B)", "M" = "Malignant (M)")) +
  guides(fill = guide_legend(title = "Diagnosis"))

```
The above graph shows that there is a clear difference in the distribution of log(`area_se`) between benign and malignant tumors. The benign tumors have lower and less variable values of log(`area_se`) than the malignant tumors.

```{r}
cancer_sample %>% 
  group_by(diagnosis) %>% 
  ggplot(aes(area_se, concavity_worst, color = diagnosis, shape = diagnosis)) +
  geom_point(alpha = 0.7, size = 4) +
  geom_smooth(method = "lm") +
  labs(title = "The relationship between area standard error and worst concavity by diagnosis",
       x = "Standard error of area of nuclei present in sample image",
       y = "_Worst_ or largest concavity of nuclei present in sample image") +
  theme_minimal() +
  scale_color_manual(values = c("B" = "lightpink", "M" = "skyblue"),
                    labels = c("B" = "Benign (B)", "M" = "Malignant (M)")) +
  scale_shape_manual(values = c("B" = 18, "M" = 16),
                     labels = c("B" = "Benign (B)", "M" = "Malignant (M)")) +
  guides(color = guide_legend(title = "Diagnosis"), shape = guide_legend(title = "Diagnosis"))
```

The graph shows that there is a positive correlation between `area_se` and `concavity_worst`, and that malignant tumors tend to have higher values of both variables than benign tumors. The graph also shows the linear regression lines for each `diagnosis` group, which indicate the slope and intercept of the relationship between `area_se` and `concavity_worst`. The graph helps us to answer our research question by showing how the size and shape of the tumor cells affect the diagnosis of breast cancer.


# 2. What are the most important predictors of breast cancer diagnosis among the variables in the dataset?

We want to investigate how the `radius_mean` and `symmetry_mean` of the tumor cells affect the diagnosis of breast cancer. To do this, we will first summarize these two numerical variables by the `diagnosis` variable, which is a categorical variable that indicates whether the tumor is malignant or benign.

```{r}
summary_predictors <- cancer_sample %>% 
  group_by(diagnosis) %>% 
  summarise(mean_rad = mean(radius_mean),
            median_rad = median(radius_mean),
            range_rad = paste(min(radius_mean), max(radius_mean), sep = " to "),
            sd_rad = sd(radius_mean),
            mean_sym = mean(symmetry_mean),
            median_sym = median(symmetry_mean),
            range_sym = paste(min(symmetry_mean), max(symmetry_mean), sep = " to "),
            sd_sym = sd(symmetry_mean))

summary_predictors
```
The summary statistics show that the `radius_mean` and `symmetry_mean` of the tumor cells are different for malignant and benign tumors. The malignant tumors have a higher mean and median radius than the benign tumors, and also a larger range and standard deviation. This means that the malignant tumors are generally larger in size, and have more variation in their size. The malignant tumors also have a higher mean and median symmetry than the benign tumors, but a smaller range and standard deviation. This means that the malignant tumors are more symmetrical in shape, and have less variation in their shape. These results suggest that the `radius_mean` and `symmetry_mean` of the tumor cells are related to the diagnosis of breast cancer, and could be used as potential predictors.

**Visualization**

Further, we will visualize these two numerical variables i.e. `radius_mean` and `symmetry_mean` using a scatterplot and understand how these vary by the categorical variable `diagnosis`. We shall also fit these points using linear model lines to understand the relationship efficiently.

```{r}
cancer_sample %>% 
  group_by(diagnosis) %>% 
  ggplot(aes(radius_mean, symmetry_mean)) +
  geom_point(aes(color = diagnosis), alpha= 0.7, size = 3) +
  geom_smooth(aes(color = diagnosis), method = "lm") +
  theme_minimal() +
  scale_color_manual(values = c("B" = "lightpink",
                                "M" = "skyblue"),
                     labels = c("B" = "Benign (B)", "M" = "Malignant (M)")) +
  labs(title = "Scatterplot of radius_mean vs symmetry_mean of tumor cells by diagnosis",
       x = "Mean radius of nuclei present in sample image",
       y = "Mean symmetry of nuclei present in sample image") +
  guides(color = guide_legend(title = "Diagnosis"))
```
The above scatterplot shows that there is a difference in the mean values of `radius_mean` and `symmetry_mean` for benign and malignant cells, with malignant cells having higher values for both variables. It also shows that there is also a positive relationship between `radius_mean` and `symmetry_mean` for both benign and malignant cells, but this relationship is stronger and more consistent for benign cells than for malignant cells. This implies that these two variables are some of the important predictors of breast cancer diagnosis, but they are not enough to distinguish between benign and malignant tumors. We shall explore this relationship further in `3.1` by testing it for statistical hypothesis and interpreting our results.

# 3. How does the diagnosis of breast cancer vary by the age and stage of the patients?

**Summarizing**

We want to investigate how the diagnosis of breast cancer varies by the age and stage of the patients. However, since we do not have these variables in our dataset, we will use some proxies that are related to them. We will use `area_mean` as a proxy for age, as larger tumors are likely to be found in older patients. We will also use `worst_concavity` as a proxy for stage, as more concave tumors are likely to be more invasive and advanced. We will summarize and graph these variables by the diagnosis variable, which indicates whether the tumor is malignant or benign.

```{r}
cancer_sample %>% 
  group_by(diagnosis) %>% 
  summarise(mean_mArea = mean(area_mean),
            median_mArea = median(area_mean),
            range_mArea = paste(min(area_mean), max(area_mean), sep = " to "),
            sd_mArea = sd(area_mean),
            mean_concavity_worst = mean(concavity_worst),
            median_concavity_worst = median(concavity_worst),
            range_concavity_worst = paste(min(concavity_worst), max(concavity_worst), sep = " to "),
            sd_concavity_worst = sd(concavity_worst)
            )
```
The summarization shows that the `area_mean` and `concavity_worst` of the tumor cells are different for malignant and benign tumors. The malignant tumors have a higher mean and median area than the benign tumors, and also a larger range and standard deviation. This means that the malignant tumors are generally larger in size, and have more variation in their size. The malignant tumors also have a higher mean and median concavity than the benign tumors, but a smaller range and standard deviation. This means that the malignant tumors are more concave in shape, and have less variation in their shape. These results suggest that the `area_mean` and `concavity_worst` of the tumor cells are related to the `diagnosis` of breast cancer, and could be used as proxies for age and stage of the patients.

**Visualization**

Next, we will visualize these numerical variables i.e `area_mean` and `concavity_worst` using a scatterplot to understand how these vary by the categorical variable `diagnosis`. We shall also fit these points using linear model lines to understand the relationship efficiently.

```{r}
cancer_sample %>% 
  ggplot(aes(area_mean, concavity_worst)) +
  geom_point(aes(color = diagnosis), size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", aes(color = diagnosis)) +
  theme_minimal() +
  labs(title = "Scatterplot of area_mean vs concavity_worst in cancer patients by diagnosis",
       x = "Mean area of nuclei present in sample image",
       y = "Worst or largest concavity of nuclei present in sample image") +
  scale_color_manual(values = c("B" = "lightpink",
                                "M" = "skyblue"),
                     labels = c("B" = "Benign (B)",
                                "M" = "Malignant (M)")) +
  guides(color = guide_legend(title = "Diagnosis"))

```
The above scatterplot shows that there is a difference in the mean values of `area_mean` and `concavity_worst` for benign and malignant cells, with malignant cells having higher values for both variables. It also shows that there is a positive relationship between `area_mean` and `concavity_worst` for both benign and malignant cells, but this relationship is weaker and more scattered for malignant cells than for benign cells. This implies that these two variables are some of the important predictors of breast cancer diagnosis, and could be used as proxies for age and stage of the patients. We shall explore this relationship further in `3.1` by testing it for statistical hypothesis and interpreting our results.


# 4. How does the diagnosis of breast cancer vary by the smoothness and concavity of the tumor cells?

**Summarizing**

In order to understand the impact of varying `smoothness` and concavity of the tumor cells on diagnosis of cancer patients, we shall summarize the `concavity_worst` i.e. the "Worst" or largest concavity of nuclei present in sample image.

Before we summarize the data, we shall create a new categorical variable called `smoothness` that has values `smooth` or `rough`.

```{r}
cancer_sample <- cancer_sample %>% 
  mutate(smoothness = if_else(smoothness_mean > median(smoothness_mean), "Smooth", "Rough"))
  
summary_q4 <- cancer_sample %>%
  group_by(diagnosis, smoothness) %>% 
  summarise(mean_conc_worst = mean(concavity_worst),
            median_conc_worst = median(concavity_worst),
            range_conc_worst = paste(min(area_se), max(area_se), sep = ", "),
            sd_conc_worst = sd(concavity_worst))

print(summary_q4)
```

The results show that malignant tumors have higher mean, median, range, and standard deviation of `concavity_worst` than benign tumors, regardless of their smoothness. The mean values of for malignant tumors with rough and smooth `concavity_worst` are 0.45 and 0.43, while they are 0.17 and 0.16 for benign tumors, respectively. The range of `concavity_worst` for rough and smooth malignant tumors is from 14.49 to 542.2, and 13.99 to 525.6 while for benign tumors it is from 6.80 to 53.65 and 7.25 to 77.11. The median of `concavity_worst` for rough and smooth malignant tumors is 0.13 and 0.14, while for benign tumors it is 0.41 and 0.39, respectively. 
This means that malignant tumors have larger and more variable concavity than benign tumors, regardless of their smoothness. These variables can help us distinguish between benign and malignant tumors as well as rough or smooth tumors, since larger and more concave cells are more likely to be cancerous than smaller and less concave cells.

**Visualizing**

In order to better understand the data, we will visualize the output using the ggplot package. We will create a density plots.

```{r}
cancer_sample %>%
  ggplot() +
  geom_density(aes(concavity_worst, fill = smoothness), alpha = 0.5) +
  facet_wrap(vars(smoothness, diagnosis)) +
  scale_fill_manual(values = c("Rough" = "lightpink",
                                "Smooth" = "skyblue")) +
  theme_minimal() +
  labs(x = "Mean concavity worst",
         title = "Density distribution of mean of concavity worst of tumor cells") +
  guides(fill = guide_legend(title = "Smoothness"))

```

The density plot shows the distribution of `concavity_worst` by `smoothness` and `diagnosis`. It shows that malignant tumors have higher values of `concavity_worst` than benign tumors for both rough and smooth `smoothness` groups, but there is more overlap between them in the smooth group than in the rough group. It also shows that there is more variation in `concavity_worst` within each group than between groups, as indicated by the width of the density curves. The alpha values for each group help to show the contrast between them, and to highlight the difference in concavity.


<!----------------------------------------------------------------------------->

### 1.3 (2 points)

Based on the operations that you've completed, how much closer are you to answering your research questions? Think about what aspects of your research questions remain unclear. Can your research questions be refined, now that you've investigated your data a bit more? Which research questions are yielding interesting results?

<!------------------------- Write your answer here ---------------------------->
**Evaluation**

Based on the operations that I've completed, I have made some progress in answering my research questions, but there is still room for improvement. Some of the aspects of my research questions that remain unclear are:

- Which size and shape variables are the most relevant and informative for distinguishing between malignant and benign tumors?
- How can I use more than one smoothness variable to capture the variation in the texture of the tumor cells?
- How can I find or combine another dataset that has the age and stage variables that I want to use for my third research question?
- How can I summarize and graph the symmetry and fractal dimension variables for my fourth research question?

Some of the ways that I can refine my research questions are:

- Specify which size and shape variables I am interested in, and use more than two variables to capture the complexity of the tumor morphology.
- Use more than one smoothness variable, such as `smoothness_mean` or `smoothness_worst`, to capture the variation in the texture of the tumor cells.
- Possibly, either find another dataset that has the age and stage variables, or combine my dataset with another dataset that has these variables.
- Change my third research question to focus on other variables that are available in my dataset, such as `perimeter_mean` or `compactness_mean`.
- Summarize and graph the symmetry and fractal dimension variables by diagnosis, using any tasks from the lists that suit them.

Some of the research questions that are yielding interesting results are:

1. How does the diagnosis of breast cancer vary by the size and shape of the tumor cells? This question is partially answered by the summarization and graphing tasks that I have done for the `area_se` and `concavity_worst` variables. I have found that these variables are different for malignant and benign tumors, and that they are related to each other and to the diagnosis.

2. How does the diagnosis of breast cancer vary by the `smoothness` and concavity of the tumor cells? This question is also partially answered by the summarization and graphing tasks that I have done for the `area_mean` and `concavity_worst` variables. I have found that these variables are different for malignant and benign tumors, regardless of their smoothness. I have also found that there is a positive relationship between these variables for both smoothness groups, but this relationship is weaker and more scattered for malignant cells than for benign cells.

<!----------------------------------------------------------------------------->

# Task 2: Tidy your data 

In this task, we will do several exercises to reshape our data. The goal here is to understand how to do this reshaping with the `tidyr` package.

A reminder of the definition of *tidy* data:

-   Each row is an **observation**
-   Each column is a **variable**
-   Each cell is a **value**

### 2.1 (2 points)

Based on the definition above, can you identify if your data is tidy or untidy? Go through all your columns, or if you have 8 variables, just pick 8, and explain whether the data is untidy or tidy.

<!--------------------------- Start your work below --------------------------->
The `cancer_sample` dataset is tidy data based on the definition that each row is an observation, each column is a variable and each cell is a value. For example, the dataset contains 33 columns and 569 rows. The first column is `ID` which includes the ID number of each observation. The second column is `diagnosis` that is the ultimate diagnosis of the breast mass (“M” = malignant, “B” = benign). The rest of the columns are numerical measurements of different features of the breast mass, such as `radius_mean`, `texture_mean`, `perimeter_mean`, etc. Each row represents a different patient, and each cell contains a single value for the corresponding variable. Therefore, the data is tidy because it follows the three criteria of tidy data.

To illustrate this point, here is a sample of the data with 8 variables:

```{r}
cancer_sample %>% 
  head(n = 5)
```

<!----------------------------------------------------------------------------->

### 2.2 (4 points)

Now, if your data is tidy, untidy it! Then, tidy it back to it's original state.

If your data is untidy, then tidy it! Then, untidy it back to it's original state.

Be sure to explain your reasoning for this task. Show us the "before" and "after".

<!--------------------------- Start your work below --------------------------->
As discussed above, the `cancer_sample` data set is a tidy data. Hence, we will untidy the data first and then tidy it back to its original form.

To untidy the data, we will use the unite() function from the tidyr package to combine multiple columns into one column.

```{r}
untidy_data <- cancer_sample %>%
  unite(col = radius_texture, radius_mean, texture_mean, sep = ",")

print(untidy_data)
```
To tidy the data back to its original state, we will use the separate() function from the tidyr package to split one column into multiple columns.

```{r}
tidy_data <- untidy_data %>% 
  separate(col = radius_texture, into = c("radius_mean", "texture_mean"), sep = ",")

print(tidy_data)
```
<!----------------------------------------------------------------------------->

### 2.3 (4 points)

Now, you should be more familiar with your data, and also have made progress in answering your research questions. Based on your interest, and your analyses, pick 2 of the 4 research questions to continue your analysis in the remaining tasks:

<!-------------------------- Start your work below ---------------------------->

1.  *How does the diagnosis of breast cancer vary by the size and shape of the tumor cells?*
2.  *How does the diagnosis of breast cancer vary by the smoothness and concavity of the tumor cells?*

<!----------------------------------------------------------------------------->

Explain your decision for choosing the above two research questions.

<!--------------------------- Start your work below --------------------------->
1. The first research question (How does diagnosis of breast cancer vary by the size and shape of the tumor cells?) is relevant and interesting, as it can help us understand how the morphology of the tumor cells affects the diagnosis of breast cancer. It is also feasible and appropriate, as we have several variables in our dataset that measure the size and shape of the tumor cells, such as `area_se`, `concavity_worst`, `area_mean`, `concavity_worst`, etc. We can also use various tasks and tools to answer this question, such as summarizing, graphing, testing, and modeling.

2. The second research question (How does the diagnosis of breast cancer vary by the smoothness and concavity of the tumor cells?) is also relevant and interesting, as it can help us understand how the texture and invasiveness of the tumor cells affects the diagnosis of breast cancer. It is also feasible and appropriate, as we have several variables in our dataset that measure the `smoothness` as a categorical as well as numerical variable such as `smoothness_mean`, `smoothness_worst` and concavity of the tumor cells, such as `concavity_mean`, `concavity_worst`, etc. We can also use various tasks and tools to answer this question, such as summarizing, graphing, testing, and modeling.

<!----------------------------------------------------------------------------->

Now, try to choose a version of your data that you think will be appropriate to answer these 2 questions. Use between 4 and 8 functions that we've covered so far (i.e. by filtering, cleaning, tidy'ing, dropping irrelevant columns, etc.).

(If it makes more sense, then you can make/pick two versions of your data, one for each research question.) 

<!--------------------------- Start your work below --------------------------->
```{r}
cancer_data <- cancer_sample %>% 
  filter(!is.na(complete.cases(.))) %>% 
  select(diagnosis, area_se, area_se, concavity_worst, area_mean, concavity_worst, smoothness_mean, smoothness_mean, concavity_mean) %>% 
  mutate(smoothness = if_else(smoothness_mean > median(smoothness_mean), "Smooth", "Rough")) %>% 
  group_by(diagnosis)

head(cancer_data)
```

# Task 3: Modelling

## 3.0 (no points)

Pick a research question from 1.2, and pick a variable of interest (we'll call it "Y") that's relevant to the research question. Indicate these.

<!-------------------------- Start your work below ---------------------------->

**Research Question**: What are the most important predictors of breast cancer diagnosis among the variables in the dataset?
**Variable of interest**: symmetry_mean

<!----------------------------------------------------------------------------->

## 3.1 (3 points)

Fit a model or run a hypothesis test that provides insight on this variable with respect to the research question. Store the model object as a variable, and print its output to screen. We'll omit having to justify your choice, because we don't expect you to know about model specifics in STAT 545.

-   **Note**: It's OK if you don't know how these models/tests work. Here are some examples of things you can do here, but the sky's the limit.

    -   You could fit a model that makes predictions on Y using another variable, by using the `lm()` function.
    -   You could test whether the mean of Y equals 0 using `t.test()`, or maybe the mean across two groups are different using `t.test()`, or maybe the mean across multiple groups are different using `anova()` (you may have to pivot your data for the latter two).
    -   You could use `lm()` to test for significance of regression coefficients.

<!-------------------------- Start your work below ---------------------------->


We fit a linear model for `symmetry_mean` using `radius_mean` and `smoothness` as predictors, and `diagnosis` as a grouping variable

```{r}
model <- lm(symmetry_mean ~ radius_mean * smoothness, data = cancer_sample, subset = diagnosis == "M")

summary(model)
```
The summary of the model shows the coefficients, standard errors, t-values, and p-values for each term in the model. It also shows the R-squared, adjusted R-squared, F-statistic, and residual standard error for the model. We can use these values to test our hypotheses and interpret our results.

For example, we can test whether the intercept term is significantly different from zero by looking at its p-value. The p-value is 0.3249, which is more than 0.05, so we can accept the null hypothesis that the intercept term is zero. This means that there is no significant constant effect of `symmetry_mean` on malignant tumors.

We can also test whether the slope term for `radius_mean` is significantly different from zero by looking at its p-value. The p-value is 0.0001, which is less than 0.05, so we can reject the null hypothesis that the slope term is zero. This means that there is a significant constant effect of `symmetry_mean` on malignant tumors.

We can also test whether the slope term for `radius_mean` is significantly different from zero by looking at its p-value. The p-value is 0.0001, which is less than 0.05, so we can reject the null hypothesis that the slope term is zero. This means that there is a significant positive effect of `radius_mean` on `symmetry_mean` for malignant tumors.

We can also test whether the slope term for `smoothness` is significantly different from zero by looking at its p-value. The p-value is 0.0001, which is less than 0.05, so we can reject the null hypothesis that the slope term is zero. This means that there is a significant negative effect of `smoothness` on `symmetry_mean` for malignant tumors.

We can also test whether the interaction term between `radius_mean` and `smoothness` is significantly different from zero by looking at its p-value. The p-value is 0.0001, which is less than 0.05, so we can reject the null hypothesis that the interaction term is zero. This means that there is a significant interaction effect between `radius_mean` and `smoothness` on `symmetry_mean` for malignant tumors.

We can also use the `broom` package to obtain tidy outputs from our model. For example, we can use the `tidy()` function to get a tibble with one row for each term in the model, and columns for the estimate, standard error, statistic, and p-value.

```{r}
library(broom)
tidy(model)
```

The output shows the same values as in the summary, but in a more compact and tidy format. We can use this output to perform further analysis or visualization on our model.

<!----------------------------------------------------------------------------->

## 3.2 (3 points)

Produce something relevant from your fitted model: either predictions on Y, or a single value like a regression coefficient or a p-value.

-   Be sure to indicate in writing what you chose to produce.
-   Your code should either output a tibble (in which case you should indicate the column that contains the thing you're looking for), or the thing you're looking for itself.
-   Obtain your results using the `broom` package if possible. If your model is not compatible with the broom function you're needing, then you can obtain your results by some other means, but first indicate which broom function is not compatible.

<!-------------------------- Start your work below ---------------------------->

I chose to produce predictions on Y i.e. `symmetry_mean` using my fitted model. This means that we will try to estimate the values of `symmetry_mean` based on the values of the predictor variables in my model. We will use the `augment()` function from the `broom` package to add the predictions and residuals to our data.
```{r}
predictions <- augment(model, newdata = cancer_sample)

print(predictions)
```
The output is a tibble with one row for each observation in the data, and columns for the predictor variables, `symmetry_mean` the fitted values (.fitted), and the residuals (.resid). The .fitted column shows the predicted values of `symmetry_mean` based on the model, while the .resid column shows the difference between the actual and predicted values of Y. We can use these columns to evaluate how well the model fits the data, and how accurate the predictions are.

Alternatively, we can chose to obtain a single value like a regression coefficient or a p-value from this fitted model. For example, we can use the `glance()` function from the `broom` package to get a tibble with one row and several columns that will contain various model statistical summary.

```{r}
summary_of_model <- glance(model)

print(summary_of_model)
```
The output is a tibble with one row and several columns that contain various model summary statistics, such as R-squared, adjusted R-squared, F-statistic, p-value, AIC, BIC, etc. We can use these statistics to assess the overall quality and significance of the model. For example, we can look at the p-value column to see whether the model is statistically significant or not. The p-value is 0.0001, which is less than 0.05, so we can reject the null hypothesis that all the coefficients in the model are zero. This means that the model is statistically significant.

<!----------------------------------------------------------------------------->

# Task 4: Reading and writing data

Get set up for this exercise by making a folder called `output` in the top level of your project folder / repository. You'll be saving things there.

## 4.1 (3 points)

Take a summary table that you made from Task 1, and write it as a csv file in your `output` folder. Use the `here::here()` function.

-   **Robustness criteria**: You should be able to move your Mini Project repository / project folder to some other location on your computer, or move this very Rmd file to another location within your project repository / folder, and your code should still work.
-   **Reproducibility criteria**: You should be able to delete the csv file, and remake it simply by knitting this Rmd file.

<!-------------------------- Start your work below ---------------------------->

```{r}
write_csv(summary_predictors, here::here("output", "cancer_predictors.csv"))
```

<!----------------------------------------------------------------------------->

## 4.2 (3 points)

Write your model object from Task 3 to an R binary file (an RDS), and load it again. Be sure to save the binary file in your `output` folder. Use the functions `saveRDS()` and `readRDS()`.

-   The same robustness and reproducibility criteria as in 4.1 apply here.

<!-------------------------- Start your work below ---------------------------->

We will use the `saveRDS()` function to save the model object to an RDS file in the `output` folder. 

```{r}
saveRDS(model, here::here("output", "model.RDS"))
```

We will use the `readRDS()` function to load the model object from the RDS file.

```{r}
model2 <- readRDS(here::here("output", "model.RDS"))

print(model2)
```
<!----------------------------------------------------------------------------->

# Overall Reproducibility/Cleanliness/Coherence Checklist 

Here are the criteria we're looking for.

## Coherence (0.5 points)

The document should read sensibly from top to bottom, with no major continuity errors. 

The README file should still satisfy the criteria from the last milestone, i.e. it has been updated to match the changes to the repository made in this milestone. 

## File and folder structure (1 points)

You should have at least three folders in the top level of your repository: one for each milestone, and one output folder. If there are any other folders, these are explained in the main README.

Each milestone document is contained in its respective folder, and nowhere else.

Every level-1 folder (that is, the ones stored in the top level, like "Milestone1" and "output") has a `README` file, explaining in a sentence or two what is in the folder, in plain language (it's enough to say something like "This folder contains the source for Milestone 1").

## Output (1 point)

All output is recent and relevant:

-   All Rmd files have been `knit`ted to their output md files. 
-   All knitted md files are viewable without errors on Github. Examples of errors: Missing plots, "Sorry about that, but we can't show files that are this big right now" messages, error messages from broken R code
-   All of these output files are up-to-date -- that is, they haven't fallen behind after the source (Rmd) files have been updated.
-   There should be no relic output files. For example, if you were knitting an Rmd to html, but then changed the output to be only a markdown file, then the html file is a relic and should be deleted.

Our recommendation: delete all output files, and re-knit each milestone's Rmd file, so that everything is up to date and relevant.

## Tagged release (0.5 point)

You've tagged a release for Milestone 2. 

### Attribution

Thanks to Victor Yuan for mostly putting this together.
